---
title: "respirometry_COMBINED"
author: "CFI"
date: "2024-06-21"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Install required packages}
require("knitr")

#install.packages("lmerTest")
library("lmerTest")
#install.packages("Matrix")
library("Matrix")
#install.packages("lme4")
library("lme4")
#install.packages("emmeans")
library("emmeans")
#install.packages("pbkrtest")
library("pbkrtest")
library("tidyverse")
library(tidyverse)
library("performance")
#install.packages("see")
library("see")
#install.packages("ggrepel")
library("ggrepel")
#install.packages("sjPlot") #for visualizing linear models
library(sjPlot)
#install.packages("MASS")
library(MASS)
#install.packages("glmmTMB") # for visualizing model checks
library(glmmTMB)
#install.packages("relaimpo") #for quantifying the effects of predictor variables
library(relaimpo)
#install.packages("outliers") #for checking out whether a data point is an outlier
library(outliers)
library(Rmisc)
#install.packages("ggeffects")
library(ggeffects)
library(forestmodel)

#install.packages("svglite")
library(svglite)

#install.packages("ggpubr")
library(ggpubr)

#install.packages("patchwork")
library(patchwork)

#turn off scientific notation
options(scipen=999)

#let dplyr print a bunch of numbers
options(dplyr.print_max = 1e9)
options(max.print = 10000)   

```


```{r Set working directory and upload data}
# resp_preprocess.data <- read.csv("C:/Users/carol/OneDrive/Documents/PhD Boston University 2019-/Astrangia_nutrient_stressors_Rotjan_lab/Data/RESP_ALL_cleaned_dead_and_broken_outlierG28Y65_deleted_SA_CORR_observed_N_A_6_21_24.csv", header=TRUE, sep=",",stringsAsFactors = FALSE)
# 
# 
# #Create a NEW column for nitrate with new categories
# 
# resp_preprocess.data <- resp_preprocess.data %>%
#   mutate(NO3_category = case_when(
#    NO3_mean_observed < 6  ~ "Low" , #need two equal signs
#      NO3_mean_observed >= 6 ~ "Elevated" 
#    )
# )
# 
# 
# resp_preprocess.data <- resp_preprocess.data %>%
#   filter(abs_resp_rate_SA_corr<0.7)
# 
# #any single datapoint that was over 4 times disparate of mean or median from rest of treatment was discarded
# #Did all pre-processing and double checked dead were deleted 6.22.24


resp_all.data <- read.csv("/Users/carolineianniello/Astrangia_Ecotox/Data/resp_combined_NO3categorized_outlier_removed_6_24_24.csv", header=TRUE, sep=",",stringsAsFactors = FALSE)

count(resp_all.data$NO3_category)
#140 high observations
#580 low observations

resp_all.data$NO3_category <- factor(resp_all.data$NO3_category, levels=c("Low", "Elevated"))
resp_all.data$sym_state<- as.factor(resp_all.data$sym_state)
resp_all.data$food<- as.factor(resp_all.data$food)
resp_all.data$temp<- as.factor(resp_all.data$temp)

```


#For Manuscript


```{r Means figure}
#create a new column "NO3_category_e_coli"
resp_all.data$NO3_category_e_coli <- paste(resp_all.data$NO3_category, resp_all.data$e_coli_plate, sep="_")

resp_all.data$NO3_category_e_coli <- factor(resp_all.data$NO3_category_e_coli , levels=c("Low_0","Low_1","Elevated_0","Elevated_1"))

# Summarize all treatments by SYM STATE and temp
sum_temp_sym <- summarySE(resp_all.data, measurevar = "abs_resp_rate_SA_corr", groupvars = c("sym_state", "temp", "NO3_category_e_coli"))

# First plot for symbiotic state
gg_temp_sym <- ggplot() +
  annotate("rect", xmin = 1.5, xmax = 2.5, ymin = -Inf, ymax = Inf, alpha = 0.2, fill = "#7e7e7e") +
  annotate("rect", xmin = 3.5, xmax = 4.5, ymin = -Inf, ymax = Inf, alpha = 0.2, fill = "#7e7e7e") +
  geom_point(data = sum_temp_sym,
    aes(x = NO3_category_e_coli,
        y = abs_resp_rate_SA_corr, col=temp, shape=sym_state, group=interaction(temp, sym_state)),
    size = 4.5,
    position = position_dodge(width = 0.8)) +
  geom_errorbar(data = sum_temp_sym, 
    aes(x = NO3_category_e_coli,
        y = abs_resp_rate_SA_corr, col=temp, 
        ymin = abs_resp_rate_SA_corr - se, ymax = abs_resp_rate_SA_corr + se,
        group=interaction(temp, sym_state)),
    width = 0.25, linewidth = 0.8,
    position = position_dodge(width = 0.8)) +
    geom_vline(xintercept = 2.5, color = "black", linetype = "solid", size = 1) +  # Black line separating NO3 categories
  theme_classic() +
  theme(legend.position = "none",  # Remove legend from the first plot
        axis.title.x = element_blank(),  # Remove x-axis label from top plot
        axis.text.x = element_blank(),  # Remove x-axis text from top plot
        axis.ticks.x = element_blank(), # Remove x-axis ticks from top plot
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 18, angle = 45, hjust = 1),
        text = element_text(family = "serif", size=20)) +
  ylim(0.06, 0.16) +
  scale_x_discrete(labels=c(
    expression("Background NO"[3]~"No E. coli"),
    expression("Background NO"[3]~"+ E. coli"),
    expression("Elevated NO"[3]~"No E. coli"),
    expression("Elevated NO"[3]~"+ E. coli")
  )) +
  labs(y = expression(paste("Absolute Rate Oxygen Consumption"))) +  # y-axis label only on the top plot
  scale_shape_discrete("Symbiotic state", labels=c("Aposymbiotic","Symbiotic")) +
  scale_color_manual("Temperature", labels=c("20°C","29°C"), values=c("dodgerblue", "darkred"))

# Summarize all treatments by FOOD and temp
sum_temp_food <- summarySE(resp_all.data, measurevar = "abs_resp_rate_SA_corr", groupvars = c("food", "temp", "NO3_category_e_coli"))

# Second plot for feeding treatment
gg_temp_food <- ggplot() +
  annotate("rect", xmin = 1.5, xmax = 2.5, ymin = -Inf, ymax = Inf, alpha = 0.2, fill = "#7e7e7e") +
  annotate("rect", xmin = 3.5, xmax = 4.5, ymin = -Inf, ymax = Inf, alpha = 0.2, fill = "#7e7e7e") +
  geom_point(data = sum_temp_food,
    aes(x = NO3_category_e_coli,
        y = abs_resp_rate_SA_corr, col=temp, shape=food, fill=temp, group=interaction(temp, food)),
    size = 4.5,
    position = position_dodge(width = 0.8)) +
  geom_errorbar(data = sum_temp_food, 
    aes(x = NO3_category_e_coli,
        y = abs_resp_rate_SA_corr, col=temp, 
        ymin = abs_resp_rate_SA_corr - se, ymax = abs_resp_rate_SA_corr + se,
        group=interaction(temp, food)),
    width = 0.25, linewidth = 0.8,
    position = position_dodge(width = 0.8)) +
    geom_vline(xintercept = 2.5, color = "black", linetype = "solid", size = 1) +  # Black line separating NO3 categories
  theme_classic() +
  theme(legend.position = "right",  # Keep legend in second plot
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),  # Remove y-axis label from bottom plot
       axis.text = element_text(size = 18, angle = 45, hjust = 1),
        text = element_text(family = "serif", size=20)) +
  ylim(0.06, 0.16) +
  scale_x_discrete(labels=c(
    expression("Background NO"[3]),
    expression("Background NO"[3]),
    expression("Elevated NO"[3]),
    expression("Elevated NO"[3])
  )) +
  labs(caption = "Data represent mean +/- SEM") +  # Caption for the second plot
  scale_shape_manual("Feeding treatment", labels = c("Fed", "Starved"), values = c(15, 25)) +
    scale_fill_manual(values = c("dodgerblue", "darkred")) +
  scale_color_manual("Temperature", labels=c("20°C","29°C"), values=c("dodgerblue", "darkred"))

# Combine the two plots, sharing the x-axis and legend, and add space between plots
gg_RESP_MEANS <- gg_temp_sym / plot_spacer() / gg_temp_food +  # Stack plots with a spacer in between
  plot_layout(heights = c(1, 0.05, 1), guides = "collect") &  # Control relative plot heights and share legend
  theme(legend.position = "right")  # Position the legend on the right

# Display the combined plot
gg_RESP_MEANS

ggsave(file="Figures/gg_RESP_MEANS.svg", plot=gg_RESP_MEANS, device='svg')
ggsave(file="Figures/gg_RESP_MEANS.png", plot=gg_RESP_MEANS)
```




#########################################
#### ARCHIVES
```{r}
#E. coli
sum_sym_temp_nitrate_EC <- summarySE(resp_EC, measurevar = "abs_resp_rate_SA_corr", groupvars = c("sym_state", "temp", "NO3_category"))
# Visualize
gg_sum_sym_temp_nitrate_EC <-ggplot(data = sum_sym_temp_nitrate_EC,
       aes(x = NO3_category,
           y = abs_resp_rate_SA_corr, col=temp, shape=sym_state)) +
    geom_point(size=4) +
    geom_errorbar(aes(ymin=abs_resp_rate_SA_corr-se, ymax=abs_resp_rate_SA_corr+se), width=0.25, linewidth=0.8) +
    ggtitle ("E. Coli Dosed Corals - Respiration Rate by Sym State and Temp") +
    theme(legend.position = "right",
          legend.title = element_text(size=12),
          axis.title = element_text(size = 10),
          legend.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 12)) +
    ylim(0.05, 0.17) +
   ##scale_x_discrete("Nitrate Level", labels=c("0-6uM","6-10uM","10-20uM","10uM"))+
    labs(x = "Nitrate Dose", 
         y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
         caption = "data represent mean +/- SEM" ) +
       # theme(legend.position="none")+
  scale_shape_discrete("Symbiotic state", labels=c("Aposymbiotic","Symbiotic")) +
    scale_color_manual("Temperature", labels=c("20°C","29°C"), values=c("dodgerblue", "darkred"))

gg_sum_sym_temp_nitrate_EC

#NO E. coli
sum_sym_temp_nitrate_NOEC <- summarySE(resp_NOEC, measurevar = "abs_resp_rate_SA_corr", groupvars = c("sym_state", "temp", "NO3_category"))
# Visualize
gg_sum_sym_temp_nitrate_NOEC <-ggplot(data = sum_sym_temp_nitrate_NOEC,
       aes(x = NO3_category,
           y = abs_resp_rate_SA_corr, col=temp, shape=sym_state)) +
    geom_point(size=4) +
    geom_errorbar(aes(ymin=abs_resp_rate_SA_corr-se, ymax=abs_resp_rate_SA_corr+se), width=0.25, linewidth=0.8) +
    ggtitle ("NO E. coli Corals - Respiration Rate by Sym State and Temp") +
    theme(legend.position = "right",
          legend.title = element_text(size=12),
          axis.title = element_text(size = 10),
          legend.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 12)) +
    ylim(0.05, 0.17) +
   ##scale_x_discrete("Nitrate Level", labels=c("0-6uM","6-10uM","10-20uM","10uM"))+
    labs(x = "Nitrate Dose", 
         y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
         caption = "data represent mean +/- SEM" ) +
       # theme(legend.position="none")+
  scale_shape_discrete("Symbiotic state", labels=c("Aposymbiotic","Symbiotic")) +
    scale_color_manual("Temperature", labels=c("20°C","29°C"), values=c("dodgerblue", "darkred"))

gg_sum_sym_temp_nitrate_NOEC
```




```{r for manuscript ARCHIVES}
resp_NOEC <- resp_all.data %>%
  filter(e_coli_plate==0) 
sum_sym_state_nitrate_NOEC <- summarySE(resp_NOEC, measurevar = "abs_resp_rate_SA_corr", groupvars = c("sym_state", "NO3_category"))
# Visualize
gg_sym_state_nitrate_NOEC <-ggplot(data = sum_sym_state_nitrate_NOEC,
       aes(x = NO3_category,
           y = abs_resp_rate_SA_corr, col=sym_state)) +
    geom_point() +
    geom_errorbar(aes(ymin=abs_resp_rate_SA_corr-se, ymax=abs_resp_rate_SA_corr+se), width=0.35, size = 1) +
    ggtitle ("Nitrate - Sym state by Dose") +
    theme(legend.position = "right",
          legend.title = element_text(size=12),
          axis.title = element_text(size = 10),
          legend.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 12)) +
    ylim(0.05, 0.25) +
   ##scale_x_discrete("Nitrate Level", labels=c("0-6uM","6-10uM","10-20uM","10uM"))+
    labs(x = "Nitrate Dose", 
         y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
         caption = "data represent mean +/- SEM" ) +
       # theme(legend.position="none")+
  scale_color_manual("Symbiotic state", labels=c("Aposymbiotic","Symbiotic"), values=c("gray", "brown"))
gg_sym_state_nitrate_NOEC


#lets t test low vs high
hist(log(resp_all.data$abs_resp_rate_SA_corr))
#looks beauty when you log it
subset_low_high <- subset(resp_all.data, NO3_category== "Low" | NO3_category== "High")
t.test(log(abs_resp_rate_SA_corr) ~ NO3_category, data=subset_low_high, paired=F) #SIGNIFICANTTTTTT FAAAK YEAH


t.test(NO3_uM ~ food, data= water_qual_bigelow_all, paired=F)


# # NO E COLI AT 20C
resp_NOEC_20C <- resp_all.data %>%
  filter(e_coli_plate==0 & temp==20)
#resp_NOEC_20C$NO3_category <- factor(resp_NOEC_20C$NO3_category, levels=c(0,5,7.5,10))
resp_NOEC_20C$sym_state<- as.factor(resp_NOEC_20C$sym_state)
sum_nitrate_NOEC_20C<- summarySE(resp_NOEC_20C, measurevar = "abs_resp_rate_SA_corr", groupvars = c("sym_state", "NO3_category"))
# Visualize
gg_nitrate_NOEC_20C <-ggplot(data = sum_nitrate_NOEC_20C,
       aes(x = NO3_category,
           y = abs_resp_rate_SA_corr, col=sym_state)) +
    geom_point() +
    geom_errorbar(aes(ymin=abs_resp_rate_SA_corr-se, ymax=abs_resp_rate_SA_corr+se), width=0.35, size = 1) +
    ggtitle ("Nitrate NO E. Coli 20C - Sym state by Dose") +
    theme(legend.position = "right",
          legend.title = element_text(size=12),
          axis.title = element_text(size = 10),
          legend.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 12)) +
   ##scale_x_discrete("Nitrate Level", labels=c("0-6uM","6-10uM","10-20uM","10uM"))+
    labs(x = "Nitrate Dose", 
         y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
         caption = "data represent mean +/- SEM" ) +
  ylim(0.05, 0.25) +
       # theme(legend.position="none")+
  scale_color_manual("Symbiotic state", labels=c("Aposymbiotic","Symbiotic"), values=c("gray", "brown"))
gg_nitrate_NOEC_20C

# # E COLI AT 20C
resp_EC_20C <- resp_all.data %>%
  filter(e_coli_plate==1 & temp==20)
resp_EC_20C$NO3_category <- factor(resp_EC_20C$NO3_category, levels=c("Low","High"))
resp_EC_20C$sym_state<- as.factor(resp_EC_20C$sym_state)
sum_nitrate_EC_20C<- summarySE(resp_EC_20C, measurevar = "abs_resp_rate_SA_corr", groupvars = c("sym_state", "NO3_category"))
# Visualize
gg_nitrate_EC_20C <-ggplot(data = sum_nitrate_EC_20C,
       aes(x = NO3_category,
           y = abs_resp_rate_SA_corr, col=sym_state)) +
    geom_point() +
    geom_errorbar(aes(ymin=abs_resp_rate_SA_corr-se, ymax=abs_resp_rate_SA_corr+se), width=0.35, size = 1) +
    ggtitle ("Nitrate E. Coli 20C - Sym state by Dose") +
    theme(legend.position = "right",
          legend.title = element_text(size=12),
          axis.title = element_text(size = 10),
          legend.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 12)) +
   #scale_x_discrete("Nitrate Level", labels=c("0-6uM","6-10uM","10-20uM","10uM"))+
    labs(x = "Nitrate Dose", 
         y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
         caption = "data represent mean +/- SEM" ) +
   ylim(0.05, 0.25) +
       # theme(legend.position="none")+
  scale_color_manual("Symbiotic state", labels=c("Aposymbiotic","Symbiotic"), values=c("gray", "brown"))
gg_nitrate_EC_20C

# # NO E COLI AT 30C
resp_NOEC_30C <- resp_all.data %>%
  filter(e_coli_plate==0 & temp==30)
resp_NOEC_30C$sym_state<- as.factor(resp_NOEC_30C$sym_state)
sum_nitrate_NOEC_30C<- summarySE(resp_NOEC_30C, measurevar = "abs_resp_rate_SA_corr", groupvars = c("sym_state", "NO3_category"))
# Visualize
gg_nitrate_NOEC_30C <-ggplot(data = sum_nitrate_NOEC_30C,
       aes(x = NO3_category,
           y = abs_resp_rate_SA_corr, col=sym_state)) +
    geom_point() +
    geom_errorbar(aes(ymin=abs_resp_rate_SA_corr-se, ymax=abs_resp_rate_SA_corr+se), width=0.35, size = 1) +
    ggtitle ("Nitrate NO E. Coli 30C - Sym state by Dose") +
    theme(legend.position = "right",
          legend.title = element_text(size=12),
          axis.title = element_text(size = 10),
          legend.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 12)) +
   #scale_x_discrete("Nitrate Level", labels=c("0-6uM","6-10uM","10-20uM","10uM"))+
    labs(x = "Nitrate Dose", 
         y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
         caption = "data represent mean +/- SEM" ) +
 ylim(0.05, 0.25) +
       # theme(legend.position="none")+
  scale_color_manual("Symbiotic state", labels=c("Aposymbiotic","Symbiotic"), values=c("gray", "brown"))
gg_nitrate_NOEC_30C

# # E COLI AT 30C
resp_EC_30C <- resp_all.data %>%
  filter(e_coli_plate==1 & temp==30)
resp_EC_30C$sym_state<- as.factor(resp_EC_30C$sym_state)
sum_nitrate_EC_30C<- summarySE(resp_EC_30C, measurevar = "abs_resp_rate_SA_corr", groupvars = c("sym_state", "NO3_category"))
# Visualize
gg_nitrate_EC_30C <-ggplot(data = sum_nitrate_EC_30C,
       aes(x = NO3_category,
           y = abs_resp_rate_SA_corr, col=sym_state)) +
    geom_point() +
    geom_errorbar(aes(ymin=abs_resp_rate_SA_corr-se, ymax=abs_resp_rate_SA_corr+se), width=0.35, size = 1) +
    ggtitle ("Nitrate E. Coli 30C - Sym state by Dose") +
    theme(legend.position = "right",
          legend.title = element_text(size=12),
          axis.title = element_text(size = 10),
          legend.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 12)) +
   #scale_x_discrete("Nitrate Level", labels=c("0-6uM","6-10uM","10-20uM","10uM"))+
    labs(x = "Nitrate Dose", 
         y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
         caption = "data represent mean +/- SEM" ) +
    ylim(0.05, 0.25) +
       # theme(legend.position="none")+
  scale_color_manual("Symbiotic state", labels=c("Aposymbiotic","Symbiotic"), values=c("gray", "brown"))
gg_nitrate_EC_30C


# # By temperature
resp_NOEC <- resp_all.data %>%
  filter(e_coli_plate==0)
resp_NOEC$sym_state<- as.factor(resp_NOEC$sym_state)
resp_NOEC$temp<- as.factor(resp_NOEC$temp)
sum_nitrate_NOEC<- summarySE(resp_NOEC, measurevar = "abs_resp_rate_SA_corr", groupvars = c("NO3_category", "temp"))
# Visualize
gg_nitrate_NOEC <-ggplot(data = sum_nitrate_NOEC,
       aes(x = NO3_category,
           y = abs_resp_rate_SA_corr, col=temp)) +
    geom_point() +
    geom_errorbar(aes(ymin=abs_resp_rate_SA_corr-se, ymax=abs_resp_rate_SA_corr+se), width=0.35, size = 1) +
    ggtitle ("Nitrate NO E. Coli - Temperature by Dose") +
    theme(legend.position = "right",
          legend.title = element_text(size=12),
          axis.title = element_text(size = 10),
          legend.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 12)) +
   #scale_x_discrete("Nitrate Level", labels=c("0-6uM","6-10uM","10-20uM","10uM"))+
    labs(x = "Nitrate Dose", 
         y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
         caption = "data represent mean +/- SEM" ) +
       # theme(legend.position="none")+
     ylim(0.05, 0.25) +
  scale_color_manual("Temperature", labels=c("20°C","29°C"), values=c("dodgerblue", "red3"))
gg_nitrate_NOEC 

# # By temp with E. coli
resp_EC <- resp_all.data %>%
  filter(e_coli_plate==1)
#resp_EC$NO3_category <- factor(resp_EC$NO3_category, levels=c(0,5,7.5,10))
resp_EC$sym_state<- as.factor(resp_EC$sym_state)
resp_EC$temp<- as.factor(resp_EC$temp)
sum_nitrate_EC<- summarySE(resp_EC, measurevar = "abs_resp_rate_SA_corr", groupvars = c("NO3_category", "temp"))
# Visualize
gg_nitrate_EC <-ggplot(data = sum_nitrate_EC,
       aes(x = NO3_category,
           y = abs_resp_rate_SA_corr, col=temp)) +
    geom_point() +
    geom_errorbar(aes(ymin=abs_resp_rate_SA_corr-se, ymax=abs_resp_rate_SA_corr+se), width=0.35, size = 1) +
    ggtitle ("Nitrate E. Coli - Temperature by Dose") +
    theme(legend.position = "right",
          legend.title = element_text(size=12),
          axis.title = element_text(size = 10),
          legend.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 12)) +
   #scale_x_discrete("Nitrate Level", labels=c("0-6uM","6-10uM","10-20uM","10uM"))+
    labs(x = "Nitrate Dose", 
         y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
         caption = "data represent mean +/- SEM" ) +
       # theme(legend.position="none")+
      ylim(0.05, 0.25) +
  scale_color_manual("Temperature", labels=c("20°C","29°C"), values=c("dodgerblue", "red3"))
gg_nitrate_EC 


############# ADDING FEEDING STATE AS A SHAPE
# # NO E COLI AT 20C
sum_nitrate_NOEC_20C_food <- summarySE(resp_NOEC_20C, measurevar = "abs_resp_rate_SA_corr", groupvars = c("sym_state", "NO3_category", "food"))
# Visualize
gg_nitrate_NOEC_20C_food <-ggplot(data = sum_nitrate_NOEC_20C_food,
       aes(x = NO3_category,
           y = abs_resp_rate_SA_corr, col=sym_state, shape=food)) +
    geom_point(size=5, position=position_dodge(width=0.5)) +
    geom_errorbar(aes(ymin=abs_resp_rate_SA_corr-se, ymax=abs_resp_rate_SA_corr+se), width=0.35, size = 1, position=position_dodge(width=0.5)) +
    ggtitle ("Nitrate NO E. Coli 20C - Sym state by Dose") +
    theme(legend.position = "right",
          legend.title = element_text(size=12),
          axis.title = element_text(size = 10),
          legend.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 12)) +
   #scale_x_discrete("Nitrate Level", labels=c("0-6uM","6-10uM","10-20uM","10uM"))+
    labs(x = "Nitrate Dose", 
         y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
         caption = "data represent mean +/- SEM" ) +
    ylim(0.05, 0.3) +
       # theme(legend.position="none")+
  scale_color_manual("Symbiotic state", labels=c("Aposymbiotic","Symbiotic"), values=c("gray", "brown"))+
scale_shape_discrete(labels=c("Fed","Starved"),name="Nutritional State")
gg_nitrate_NOEC_20C_food

# # E COLI AT 20C
sum_nitrate_EC_20C_food<- summarySE(resp_EC_20C, measurevar = "abs_resp_rate_SA_corr", groupvars = c("sym_state", "NO3_category","food"))
# Visualize
gg_nitrate_EC_20C_food <-ggplot(data = sum_nitrate_EC_20C_food,
       aes(x = NO3_category,
           y = abs_resp_rate_SA_corr, col=sym_state, shape=food)) +
   geom_point(size=5, position=position_dodge(width=0.5)) +
    geom_errorbar(aes(ymin=abs_resp_rate_SA_corr-se, ymax=abs_resp_rate_SA_corr+se), width=0.35, size = 1, position=position_dodge(width=0.5)) +
    ggtitle ("Nitrate E. Coli 20C - Sym state by Dose") +
    theme(legend.position = "right",
          legend.title = element_text(size=12),
          axis.title = element_text(size = 10),
          legend.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 12)) +
   #scale_x_discrete("Nitrate Level", labels=c("0-6uM","6-10uM","10-20uM","10uM"))+
    labs(x = "Nitrate Dose", 
         y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
         caption = "data represent mean +/- SEM" ) +
 ylim(0.05, 0.3) +
       # theme(legend.position="none")+
  scale_color_manual("Symbiotic state", labels=c("Aposymbiotic","Symbiotic"), values=c("gray", "brown"))+
scale_shape_discrete(labels=c("Fed","Starved"),name="Nutritional State")
gg_nitrate_EC_20C_food

# # NO E COLI AT 30C
sum_nitrate_NOEC_30C_food<- summarySE(resp_NOEC_30C, measurevar = "abs_resp_rate_SA_corr", groupvars = c("sym_state", "NO3_category","food"), na.rm=T)
# Visualize
gg_nitrate_NOEC_30C_food <-ggplot(data = sum_nitrate_NOEC_30C_food,
       aes(x = NO3_category,
           y = abs_resp_rate_SA_corr, col=sym_state, shape=food)) +
   geom_point(size=5, position=position_dodge(width=0.5)) +
    geom_errorbar(aes(ymin=abs_resp_rate_SA_corr-se, ymax=abs_resp_rate_SA_corr+se), width=0.35, size = 1, position=position_dodge(width=0.5)) +
    ggtitle ("Nitrate NO E. Coli 30C - Sym state by Dose") +
    theme(legend.position = "right",
          legend.title = element_text(size=12),
          axis.title = element_text(size = 10),
          legend.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 12)) +
   #scale_x_discrete("Nitrate Level", labels=c("0-6uM","6-10uM","10-20uM","10uM"))+
    labs(x = "Nitrate Dose", 
         y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
         caption = "data represent mean +/- SEM" ) +
  ylim(0.05, 0.3) +
       # theme(legend.position="none")+
  scale_color_manual("Symbiotic state", labels=c("Aposymbiotic","Symbiotic"), values=c("gray", "brown"))+
scale_shape_discrete(labels=c("Fed","Starved"),name="Nutritional State")
gg_nitrate_NOEC_30C_food

# # E COLI AT 30C
sum_nitrate_EC_30C_food<- summarySE(resp_EC_30C, measurevar = "abs_resp_rate_SA_corr", groupvars = c("sym_state", "NO3_category", "food"))
# Visualize
gg_nitrate_EC_30C_food <-ggplot(data = sum_nitrate_EC_30C_food,
       aes(x = NO3_category,
           y = abs_resp_rate_SA_corr, col=sym_state, shape=food)) +
    geom_point(size=5, position=position_dodge(width=0.5)) +
    geom_errorbar(aes(ymin=abs_resp_rate_SA_corr-se, ymax=abs_resp_rate_SA_corr+se), width=0.35, size = 1, position=position_dodge(width=0.5)) +
    ggtitle ("Nitrate E. Coli 30C - Sym state by Dose") +
    theme(legend.position = "right",
          legend.title = element_text(size=12),
          axis.title = element_text(size = 10),
          legend.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 12)) +
   #scale_x_discrete("Nitrate Level", labels=c("0-6uM","6-10uM","10-20uM","10uM"))+
    labs(x = "Nitrate Dose", 
         y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
         caption = "data represent mean +/- SEM" ) +
  ylim(0.05, 0.3) +
       # theme(legend.position="none")+
  scale_color_manual("Symbiotic state", labels=c("Aposymbiotic","Symbiotic"), values=c("gray", "brown"))+
scale_shape_discrete(labels=c("Fed","Starved"),name="Nutritional State")
gg_nitrate_EC_30C_food


# # By temperature
sum_nitrate_NOEC_food<- summarySE(resp_NOEC, measurevar = "abs_resp_rate_SA_corr", groupvars = c("NO3_category", "temp", "food"))
# Visualize
gg_nitrate_NOEC_food <-ggplot(data = sum_nitrate_NOEC_food,
       aes(x = NO3_category,
           y = abs_resp_rate_SA_corr, col=temp, shape=food)) +
    geom_point(size=5, position=position_dodge(width=0.5)) +
    geom_errorbar(aes(ymin=abs_resp_rate_SA_corr-se, ymax=abs_resp_rate_SA_corr+se), width=0.35, size = 1, position=position_dodge(width=0.5)) +
    ggtitle ("Nitrate NO E. Coli - Temperature by Dose") +
    theme(legend.position = "right",
          legend.title = element_text(size=12),
          axis.title = element_text(size = 10),
          legend.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 12)) +
   #scale_x_discrete("Nitrate Level", labels=c("0-6uM","6-10uM","10-20uM","10uM"))+
    labs(x = "Nitrate Dose", 
         y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
         caption = "data represent mean +/- SEM" ) +
  scale_shape_discrete(labels=c("Fed","Starved"),name="Nutritional State") +
       # theme(legend.position="none")+
  ylim(0.05, 0.3) +
  scale_color_manual("Temperature", labels=c("20°C","29°C"), values=c("dodgerblue", "red3"))
gg_nitrate_NOEC_food

# # By temp with E. coli
sum_nitrate_EC_food<- summarySE(resp_EC, measurevar = "abs_resp_rate_SA_corr", groupvars = c("NO3_category", "temp","food"))
# Visualize
gg_nitrate_EC <-ggplot(data = sum_nitrate_EC_food,
       aes(x = NO3_category,
           y = abs_resp_rate_SA_corr, col=temp, shape=food)) +
    geom_point(size=5, position=position_dodge(width=0.5)) +
    geom_errorbar(aes(ymin=abs_resp_rate_SA_corr-se, ymax=abs_resp_rate_SA_corr+se), width=0.35, size = 1, position=position_dodge(width=0.5)) +
    ggtitle ("Nitrate E. Coli - Temperature by Dose") +
    theme(legend.position = "right",
          legend.title = element_text(size=12),
          axis.title = element_text(size = 10),
          legend.background = element_blank(),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 12)) +
   #scale_x_discrete("Nitrate Level", labels=c("0-6uM","6-10uM","10-20uM","10uM"))+
    labs(x = "Nitrate Dose", 
         y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
         caption = "data represent mean +/- SEM" ) +
       # theme(legend.position="none")+
  ylim(0.05, 0.3) +
  scale_shape_discrete(labels=c("Fed","Starved"),name="Nutritional State") +
  scale_color_manual("Temperature", labels=c("20°C","29°C"), values=c("dodgerblue", "red3"))
gg_nitrate_EC 
```

```{r Generating subsets to work with}


#SUBSET - No E. coli
resp_NOEC <- resp_all.data %>%
  filter(e_coli_plate==0)
#SUBSET - E. coli
resp_EC <- resp_all.data %>%
  filter(e_coli_plate==1)


#SUBSET - NO E. coli 20°C
resp_NOEC_20 <- resp_all.data %>%
  filter(e_coli_plate==0) %>%
  filter(temp==20)
#SUBSET - NO E. coli 29°C
resp_NOEC_30 <- resp_all.data %>%
  filter(e_coli_plate==0)%>%
  filter(temp==30)

#SUBSET - E. coli 20°C
resp_EC_20 <- resp_all.data %>%
  filter(e_coli_plate==1) %>%
  filter(temp==20)
#SUBSET - E. coli 29°C
resp_EC_30 <- resp_all.data %>%
  filter(e_coli_plate==1)%>%
  filter(temp==30)



```


```{r T tests}
#selected t tests

# NO E COLI
#T test low vs. high, no EC
t.test(abs_resp_rate_SA_corr ~ NO3_category, data= resp_NOEC)
#No overall difference in low vs. elevated nitate in respiration WITHOUT E. coli

t.test(abs_resp_rate_SA_corr ~ NO3_category, data= resp_NOEC_20)
#The above trend persists at 20C

t.test(abs_resp_rate_SA_corr ~ NO3_category, data= resp_NOEC_30)
#The above trend persists at 30C as well 

#Take home-- no differences in resp rate under no E. coli

##########
# E COLI
#T test low vs. high, no EC
t.test(abs_resp_rate_SA_corr ~ NO3_category, data= resp_EC)
#Corals under elevated nitrate respire MORE

t.test(abs_resp_rate_SA_corr ~ NO3_category, data= resp_EC_20)
#The above trend persists at 20C

t.test(abs_resp_rate_SA_corr ~ NO3_category, data= resp_EC_30)
#The above trend BREAKS DOWN at 30C!

# Take home -- Corals under E. coli treatment exhibit increased respiration at 20C, but this trend breaks down at 30C -- perhaps they are unable to mount the appropriate metabolic immune response


# SYM STATE -- sym corals respire less than apo at 30C -- parisitsm!?!
t.test(abs_resp_rate_SA_corr ~ sym_state, data= resp_NOEC)
#not significantly different
t.test(abs_resp_rate_SA_corr ~ sym_state, data= resp_EC)
#SIGNIFICANTLY DIFFERENT -- symbiotic corals respired LESS than apo
t.test(abs_resp_rate_SA_corr ~ sym_state, data= resp_EC_20)
#not significantly different
t.test(abs_resp_rate_SA_corr ~ sym_state, data= resp_EC_30)
#SIGNIFICANTLY DIFFERENT -- symbiotic corals respired LESS than apo SPECIFICALLY AT 30C!

# FEEDING --Nothing really going on
t.test(abs_resp_rate_SA_corr ~ food, data= resp_NOEC)
#not significantly different
t.test(abs_resp_rate_SA_corr ~ food, data= resp_EC)
#not significantly different
t.test(abs_resp_rate_SA_corr ~ food, data= resp_EC_20)
#not significantly different
t.test(abs_resp_rate_SA_corr ~ food, data= resp_EC_30)
#not significantly different

```

```{r ANOVAs to match graphs}
resp_EC$log_resp_rate <- log(resp_EC$abs_resp_rate_SA_corr)
hist(resp_EC$log_resp_rate)

#Now an ANOVA on sym state, temp, and NO3_category 
anova_EC <- aov(log_resp_rate ~ sym_state*temp*NO3_category, data=resp_EC)
summary(anova_EC)
TukeyHSD(anova_EC)

#without E. coli now
resp_NOEC$log_resp_rate <- log(resp_NOEC$abs_resp_rate_SA_corr)
hist(resp_NOEC$log_resp_rate)

#Now an ANOVA on sym state, temp, and NO3_category 
anova_NOEC <- aov(log_resp_rate ~ sym_state*temp*NO3_category, data=resp_NOEC)
summary(anova_NOEC)
TukeyHSD(anova_NOEC)
```


```{r}
```{r}
#create a new column "NO3_category_e_coli"
resp_all.data$NO3_category_e_coli <- paste(resp_all.data$NO3_category, resp_all.data$e_coli_plate, sep="_")

resp_all.data$NO3_category_e_coli <- factor(resp_all.data$NO3_category_e_coli , levels=c("Low_0","Low_1","Elevated_0","Elevated_1"))

#Summarize all treatments by SYM STATE and temp
sum_temp_sym <- summarySE(resp_all.data, measurevar = "abs_resp_rate_SA_corr", groupvars = c("sym_state", "temp", "NO3_category_e_coli"))
gg_temp_sym <- ggplot() +
  # Add shaded background for specific x-axis categories using geom_rect
  annotate("rect", xmin = 1.5, xmax = 2.5, ymin = -Inf, ymax = Inf, alpha = 0.2, fill = "#7e7e7e") +
  annotate("rect", xmin = 3.5, xmax = 4.5, ymin = -Inf, ymax = Inf, alpha = 0.2, fill = "#7e7e7e") +
  geom_point(data = sum_temp_sym,
    aes(x = NO3_category_e_coli,
        y = abs_resp_rate_SA_corr, col=temp, shape=sym_state, group=interaction(temp, sym_state)),
    size = 4.5,
    position = position_dodge(width = 0.8)) +
  geom_errorbar(data = sum_temp_sym, 
    aes(x = NO3_category_e_coli,
        y = abs_resp_rate_SA_corr, col=temp, 
        ymin = abs_resp_rate_SA_corr - se, ymax = abs_resp_rate_SA_corr + se,
        group=interaction(temp, sym_state)),
    width = 0.25, linewidth = 0.8,
    position = position_dodge(width = 0.8)) +
  ggtitle("Respiration Rate By Treatment, Temp, and Food") +
  theme_classic() +
  theme(legend.position = "right",
        legend.text = element_text(size=13),
        legend.title = element_text(size=15),
        axis.title = element_text(size = 14),
        legend.background = element_blank(),
        axis.text = element_text(size = 14, angle = 45, hjust = 1),
        text = element_text(family = "serif")) +
  ylim(0.06, 0.16) +
  # Updating the x-axis labels with subscripts
  scale_x_discrete("Treatment", labels=c(
    expression("Background NO"[3]~"No E. coli"),
    expression("Background NO"[3]~"+ E. coli"),
    expression("Elevated NO"[3]~"No E. coli"),
    expression("Elevated NO"[3]~"+ E. coli")
  )) +
  labs(x = "Treatment", 
       y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
       caption = "data represent mean +/- SEM") +
  scale_shape_discrete("Symbiotic state", labels=c("Aposymbiotic","Symbiotic")) +
  scale_color_manual("Temperature", labels=c("20°C","29°C"), values=c("dodgerblue", "darkred")) +
  theme(aspect.ratio = 1.2)
gg_temp_sym

#Summarize all treatments by FOOD and temp
sum_temp_food <- summarySE(resp_all.data, measurevar = "abs_resp_rate_SA_corr", groupvars = c("food", "temp", "NO3_category_e_coli"))

gg_temp_food <- ggplot() +
  # Add shaded background for specific x-axis categories using geom_rect
  annotate("rect", xmin = 1.5, xmax = 2.5, ymin = -Inf, ymax = Inf, alpha = 0.2, fill = "#7e7e7e") +
  annotate("rect", xmin = 3.5, xmax = 4.5, ymin = -Inf, ymax = Inf, alpha = 0.2, fill = "#7e7e7e") +
  geom_point(data = sum_temp_food,
    aes(x = NO3_category_e_coli,
        y = abs_resp_rate_SA_corr, col=temp, shape=food, group=interaction(temp, food)),
    size = 4.5,
    position = position_dodge(width = 0.8)) +
  geom_errorbar(data = sum_temp_food, 
    aes(x = NO3_category_e_coli,
        y = abs_resp_rate_SA_corr, col=temp, 
        ymin = abs_resp_rate_SA_corr - se, ymax = abs_resp_rate_SA_corr + se,
        group=interaction(temp, food)),
    width = 0.25, linewidth = 0.8,
    position = position_dodge(width = 0.8)) +
  ggtitle("Respiration Rate By Treatment, Temp, and Food") +
  theme_classic() +
  theme(legend.position = "right",
        legend.text = element_text(size=13),
        legend.title = element_text(size=15),
        axis.title = element_text(size = 14),
        legend.background = element_blank(),
        axis.text = element_text(size = 14, angle = 45, hjust = 1),
        text = element_text(family = "serif")) +
  ylim(0.06, 0.16) +
  # Updating the x-axis labels with subscripts
  scale_x_discrete("Treatment", labels=c(
    expression("Background NO"[3]~"No E. coli"),
    expression("Background NO"[3]~"+ E. coli"),
    expression("Elevated NO"[3]~"No E. coli"),
    expression("Elevated NO"[3]~"+ E. coli")
  )) +
  labs(x = "Treatment", 
       y = expression(paste("Absolute Rate O2 Consumption (umol*min"^"-1"~"*cm"^"2"~")")), 
       caption = "data represent mean +/- SEM") +
  scale_shape_manual("Feeding treatment", labels=c("Fed","Starved"), values=c(15,8)) +
  scale_color_manual("Temperature", labels=c("20°C","29°C"), values=c("dodgerblue", "darkred")) +
  theme(aspect.ratio = 1.2)
gg_temp_food

#NOW LETS STACK THE PLOTS
gg_RESP_MEANS <- ggarrange(gg_temp_sym, gg_temp_food, nrow=2)
gg_RESP_MEANS
```

